---
title: "Practice Assignment 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load in the libraries and the dataset

Get the family size by adding "# of siblings/spouses" and "# of parents/children" aboard the Titanic.

```{r}
#install.packages("ggplot2")
library(ggplot2)    # using "ggplot2" package to plot graphs

titanic <- read.csv("titanic.csv")
titanic$family_size <- titanic$SibSp + titanic$Parch + 1
```

## 1. Generate a series of bar charts to describe

### a. Gender of the passengers onboard

```{r}
ggplot(titanic, aes(x=Sex, fill=Sex)) + 
  geom_bar() + 
  labs(title = "Gender of the passengers onboard")
```

### b. Ticket class of the passengers onboard

```{r}
library(dplyr)      # using "dplyr" package to plot graphs
titanic %>%
  mutate(Pclass = factor(x=Pclass, levels=c(1,2,3), labels=c("1st","2nd","3rd"))) %>%
  ggplot(aes(x=Pclass, fill=Pclass)) + 
    geom_bar() + 
    ggtitle("Ticket class of the passengers onboard") + 
    xlab("Ticket Class") +
    theme(legend.position="none")
```

### c. Survival of the passengers onboard

```{r} 
titanic %>%
  mutate(Survived = factor(x=Survived, levels=c(0,1), labels=c("No","Yes"))) %>%
  ggplot(aes(x=Survived, fill=Survived)) + 
    geom_bar() + 
    ggtitle("Survival of the passengers onboard") + 
    theme(legend.position="none")
```

## 2. Generate graphs for the passengers' age

### a. Histogram 

```{r}
ggplot(titanic, aes(x=Age, y=..density..)) + 
  geom_histogram(fill="steelblue", bins=10, na.rm = TRUE) + 
  ggtitle("Distribution of age of the passengers onboard") 
```

### b. Boxplot (group by ticket class)

```{r}
titanic %>%
  mutate(Pclass = factor(x=Pclass, levels=c(1,2,3), labels=c("1st","2nd","3rd"))) %>%
  ggplot(aes(x=Pclass, y=Age, fill=Pclass)) + 
    geom_boxplot(na.rm=TRUE) + 
    ggtitle("Age of the passengers onboard") + 
    xlab("Ticket Class") + 
    theme(legend.position="none")
```

### c. Boxplot (group by survival)

```{r}
titanic %>%
  mutate(Survived = factor(x=Survived, levels=c(0,1), labels=c("No","Yes"))) %>%
  mutate(Pclass = factor(x=Pclass, levels=c(1,2,3), labels=c("1st","2nd","3rd"))) %>%
  ggplot(aes(x=Survived, y=Age, fill=Survived)) + 
    geom_boxplot(na.rm=TRUE) + 
    ggtitle("Age of the passengers onboard") + 
    theme(legend.position="none")
```

## 3. Generate a histogram for the travel fare

```{r}
ggplot(titanic, aes(x=Fare, y=..density..)) + 
  geom_histogram(fill="steelblue", bins=30, na.rm=TRUE) + 
  ggtitle("Ticket fare of the passengers onboard") 
```

Table showing the number of people who did not pay.

```{r}
paid_table <- table(titanic$Fare != 0)
names(paid_table) <- c("not pay","paid")
paid_table
```

## 4. A chart to describe the family size per ticket class

```{r}
titanic %>%
  mutate(Pclass = factor(x=Pclass, levels=c(1,2,3), labels=c("1st","2nd","3rd"))) %>%
  ggplot(aes(x=family_size, y=..density.., fill=Pclass)) +
  geom_histogram(bins=11) +
  facet_grid(Pclass ~ ., scales="free")  + 
  guides(fill=FALSE) +
  ggtitle("Family size per ticket class")
```

## 5. A series of stacked bar charts to show how survival differs

### a. Survival rate for different gender

```{r}
titanic %>%
    mutate(Survived = factor(x=Survived, levels=c(0,1), labels=c("No","Yes"))) %>%
    ggplot(aes(fill=Survived, x=Sex)) + 
    geom_bar(position="stack") +
    ggtitle("Survival by gender") + 
    xlab("Gender") + 
    guides(fill=guide_legend("Survival"))
```

### b. Survival rate for different ticket class

```{r}
titanic %>%
    mutate(Survived = factor(x=Survived, levels=c(0,1), labels=c("No","Yes"))) %>%
    mutate(Pclass = factor(x=Pclass, levels=c(1,2,3), labels=c("1st","2nd","3rd"))) %>%
    ggplot(aes(fill=Survived, x=Pclass)) + 
    geom_bar(position="stack") +
    ggtitle("Survival by ticket class") + 
    xlab("Ticket Class") + 
    guides(fill=guide_legend("Survival"))
```

## 6. A violin chart describing how survival related to age and gender 

```{r}
titanic %>%
  mutate(Survived = factor(x=Survived, levels=c(0,1), labels=c("No","Yes"))) %>%
  ggplot(aes(x=Sex, y=Age, fill=Survived)) + 
    geom_violin(na.rm=TRUE, adjust=0.5)
```

## 7. A violin chart describing the survival rate related to age and ticket class

```{r}
titanic %>%
  mutate(Survived = factor(x=Survived, levels=c(0,1), labels=c("No","Yes"))) %>%
  mutate(Pclass = factor(x=Pclass, levels=c(1,2,3), labels=c("1st","2nd","3rd"))) %>%
  ggplot(aes(x=Pclass, y=Age, fill=Survived)) + 
    geom_violin(na.rm=TRUE, adjust=0.5) + 
    xlab("Ticket Class")
```
